Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-04-06 ms.tex
4

The minimum heating rate for thermal stability (eq.~[\ref{eq:cooling2}]) corresponds to maximum thermally-stable accretion rate,
\begin{align}
\frac{\dot{M}_{\rm TI}}{\dot{M}_{\rm edd}} \simeq
  {\rm min}\begin{cases}
   5\times 10^{-4} \eta_{0.02}^{0.47} M_{\bullet,8}^{0.36} &, \text{core}, \zeta > \zeta_c\\
3\times 10^{-6}\eta_{0.02}M_{\bullet,8}^{-0.38} &, \text{core}, \zeta < \zeta_c \\
   2\times 10^{-4} \eta_{0.02}^{0.59} M_{\bullet,8}^{0.28}    &, \text{cusp}, \zeta > \zeta_c\\
2.5\times 10^{-4}\eta_{0.02}&, \text{cusp}, \zeta < \zeta_c
  \end{cases}
  \label{eq:Mdotmax}
\end{align}
where the first line for each case is derived using equation
(\ref{eq:eddr_analytic}) to calculate $\dot{M}(v_{\rm TI})$, whereas
the second line corresponds to the low wind heating limit $\zeta \ll \zeta_c$.  $\zeta_c$ is
calculated using the empirical correlation for core galaxies
$r_{b}/r_{\rm soi} \approx 7.6M_{\bullet,8}^{-0.2}$ which follows from
equation (\ref{eq:rsoi}) and the $r_b-M_{\bullet}$ correlation derived from the results of \citet{LauerFaber+:2007a}.

@@

Our analytic estimates accurately reproduce the numerical results in
the high heating limit $\zeta \gg 1$ ($v_w \gg \sigma_0$; $r_{\rm s}
\lesssim r_{\rm soi}$).  However, for sufficiently low heating the
stagnation radius diverges above the analytic estimate, approaching
the stellar break radius $r_b \gg r_{\rm soi}$.  This divergence
occurs approximately when $\zeta < \zeta_c \propto (r_b/r_{\rm
  soi})^{0.5(\Gamma-1)}$ (eq.~[\ref{eq:zetacrit}]), which in Appendix
\ref{app:be} is shown to coincide with the Bernoulli parameter of the gas decreasing below unity near the stagnation radius.  This divergence of the stagnation radius to large radii hence occurs at a higher value of $\zeta$ in core galaxies ($\Gamma = 0.8$) with large break radii.  This dependence explains the behavior of the two core galaxies, shown as
orange triangles vertically aligned at the same value of
$\zeta$. These correspond to models with the same black hole mass
($\Mbh = 10^8 \Msun$) and heating rate ($v_w = 600$ km s$^{-1}$) but
different values of the break radius: $r_{b} =$ 25, 50 pc (from top to
bottom). {\bf AG still needs work.  BDM: what work? this looks sufficient to me}

@@

In addition to our standard set of solutions neglecting thermal conduction and cooling, we explore a few test simulations including these effects, as marked by dagger symbols in Table \ref{table:models}.  For solutions which are far from being thermally unstable when cooling is neglected (e.g. $M_{\bullet} = 10^{8}M_{\odot}$, $v_{w} = 600$ km s$^{-1}$, $r_{b} = 100$ pc, $\eta = 0.2$, $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}| = 76$), the effects of including radiative cooling on the key properties of the flow (stagnation radius, mass accretion rate, $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$ ratio) are small.  However, for solutions in which cooling stability is marginal (e.g. $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}| \sim $few), including cooling decreases the value of $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$ by a greater fraction.  In particular, we ran solutions with radiative cooling for all cusp galaxies with $v_w=300$ km s$^{-1}$ and all core galaxies with $v_w=600 $ km s$^{-1}$, each for three different values of the mass input parameter $\eta = 0.02,
0.2, 0.6$, spanning the physical range.  Solutions found to be unstable for all values of $\eta$ are marked with a $\mathbf{TI}$ in Table \ref{table:models}.

We find that the key properties of solutions including thermal conduction are always minor for our choice of the saturation parameter $\phi = 0.1$, and they act to offset the effects of radiative cooling, i.e. to make the flow more thermally stable (e.g.~\citealt{Zakamska&Narayan03}).

>>>>

The minimum heating rate for thermal stability (eq.~[\ref{eq:cooling2}]) corresponds to maximum thermally-stable accretion rate,
\begin{align}
\frac{\dot{M}_{\rm TI}}{\dot{M}_{\rm edd}} \simeq
  {\rm min}\begin{cases}
   5\times 10^{-4} \eta_{0.02}^{0.47} M_{\bullet,8}^{0.36} &, \text{core}, \zeta > \zeta_c\\
3\times 10^{-6}\eta_{0.02}M_{\bullet,8}^{-0.38} &, \text{core}, \zeta < \zeta_c \\
   2\times 10^{-4} \eta_{0.02}^{0.59} M_{\bullet,8}^{0.28}    &, \text{cusp}, \zeta > \zeta_c\\
2.5\times 10^{-4}\eta_{0.02}&, \text{cusp}, \zeta < \zeta_c
  \end{cases}
  \label{eq:Mdotmax}
\end{align}
where the first line for each case is derived using equation
(\ref{eq:eddr_analytic}) to calculate $\dot{M}(v_{\rm TI})$, whereas
the second line corresponds to the low wind heating limit $\zeta \ll \zeta_c$.  $\zeta_c$ is
calculated using the empirical correlation for core galaxies
$r_{b}/r_{\rm soi} \approx 7.6M_{\bullet,8}^{0.2}$ which follows from
equation (\ref{eq:rsoi}) and the $r_b-M_{\bullet}$ correlation derived from the results of \citet{LauerFaber+:2007a}.  {\bf BDM: check this}

@@

Our analytic estimates accurately reproduce the numerical results in
the high heating limit $\zeta \gg 1$ ($v_w \gg \sigma_0$; $r_{\rm s}
\lesssim r_{\rm soi}$).  However, for sufficiently low heating the
stagnation radius diverges above the analytic estimate, approaching
the stellar break radius $r_b \gg r_{\rm soi}$.  This divergence
occurs approximately when $\zeta < \zeta_c \propto (r_b/r_{\rm
  soi})^{0.5(\Gamma-1)}$ (eq.~[\ref{eq:zetacrit}]), which in Appendix
\ref{app:be} is shown to coincide with the Bernoulli parameter of the gas decreasing below unity near the stagnation radius.  This divergence of the stagnation radius to large radii hence occurs at a higher value of $\zeta$ in core galaxies ($\Gamma = 0.8$) with large break radii.  This dependence explains the behavior of the two core galaxies, shown as
orange triangles vertically aligned at the same value of
$\zeta$; the solutions correspond to the same black hole mass
($\Mbh = 10^8 \Msun$) and heating rate ($v_w = 600$ km s$^{-1}$) but are calculated assuming
different values of the break radius: $r_{b} =$ 25, 50 pc (from top to
bottom).

@@

In addition to our standard set of solutions which neglect thermal conduction and radiative cooling, we explore a few test simulations including these effects, as marked by dagger symbols in Table \ref{table:models}.  For solutions which are far from being thermally unstable when cooling is neglected (e.g., $M_{\bullet} = 10^{8}M_{\odot}$, $v_{w} = 600$ km s$^{-1}$, $r_{b} = 100$ pc, $\eta = 0.2$, $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}| = 76$), including radiative cooling has little effect on the key properties of the flow (e.g., stagnation radius, mass accretion rate, $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$).  However, for solutions which are marginally stable (e.g., $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}| \sim $few), the inclusion of radiative acts to decrease the value of $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$ by a greater amount.  We ran solutions with radiative cooling for all cusp galaxies with $v_w=300$ km s$^{-1}$ and all core galaxies with $v_w=600 $ km s$^{-1}$, each for three different values of the mass input parameter $\eta = 0.02, 0.2, 0.6$, spanning the physical range.  Solutions found to be unstable for all values of $\eta$ are marked in boldface as ``TI" in Table \ref{table:models}.
By contrast, changes due to the inclusion of thermal conduction are always minor for our choice of the saturation parameter $\phi = 0.1$, although conduction is seen to offset the effects of radiative cooling, making the flow more thermally stable (e.g.~\citealt{Zakamska&Narayan03}).
